---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

For this assignment we were tasked with examining Seattle Public Library (SPL) data. The specific .csv data file I chose to analyze was collected by the SPL "includ[ing] a monthly count of Seattle Public Library checkouts by title for physical and electronic items" spanning from 2013 to 2023. Using this data set I specifically decided to take a look at the effect COVID-19 had on SPL checkouts. I examined the various different types of medias that can be checked out from the library; how the check out rate has fluctuated for each media over the past 10 years -- observing drastic changes in recent times during the COVID-19 pandemic. Along with that, I  more specifically took a closer look at the year 2020; the defining year of the pandemic. For the year 2020, I analyzed more specific statistics including the leading creators at the SPL in 2020, subjects, and more. 

I chose to focus on 2020 and the effects of COVID-19 in relation to the SPL because I was compelled to learn about more local impacts of the pandemic. Everything we saw on the news in the year 2020 was about COVID-19 cases and deaths, people having to work from home, and school cancellations. The effects the pandemic had on library checkouts is not exactly the first thing I think of when I think about the effects of the pandemic, so i was excited to learn more.

### Summary Information

Using the data set, I was able to analyze the drastic effect COVID-19 had on how many people were checking out different medias at the Seattle Public Library. As can be seen in the "trends over time" graph, in the year 2020 book checkouts took a massive drop. In 2019 the lowest number of books that had been checkout in a month was around 190,000 while the lowest number of books that had been checked out in a month in 2020 was 641. That is an insurmountable decrease. In 2020, it can be observed that eBooks and audio books continue on the rise. Putting this increase for eBooks and audio books into perspective, around 600,000 books were checked out in the year 2022, while around 1,000,000 eBooks were checked out in 2022. In terms of the increase in audio book check outs, it can be observed that on average an audio book was checked out 21 times in 2022, while on average a book was checked out 10 times in 2022. It is crazy to that this number is more than double.

Additionally, I looked at trends in terms of the most popular subjects and creators for items checked out. As can be seen in my third graph, the most popular subject checked out in 2022 was fiction with 897,379 item check outs. This did not particularly surprise me as I also computed the top authors of 2020; being ____, who is a fictional writer. 







### The Dataset

The Seattle Public Library data set that I am analyzing for the purposes of this assignment can be found on the "Seattle Open Data" program website. The Open Data program makes it possible for data collected and generated by the City of Seattle to be publicly available. This specific data set was collected from multiple sources. In terms of Digital Items, media vendors: Overdrive, hoopla, Freegal, and RBDigital are the source of the provided usage data. For physical "item checkouts from April 2005 to September 30, 2016, the source of this data is the Legrady artwork data archives[, and for] For October 1 2016 to present, the source is the Horizon ILS." One can assume that this data was collected to keep track of item/material inventory or generate an economic opportunity for those that may benefit from the knowledge created the data. Though, there was no clear indicator as to why this data set was made publicly available.

When working with data sets as large as this, it can become important to filter and narrow down the specific data you are looking for. The data set I was working with spanned about 10 years and had over 4 million entries, each comprised of 12 columns. Additionally, it must be kept in mind when looking through my graphs and calculations that the specific data set I looked at only included items that were checked out at least 5 times a month.

Sorting through this data was difficult as not all of the information was inputted the same, or spelled correctly. When filtering through the book subjects there were more that 225,000 different subjects I had to narrow down, and when looking through some of the book titles it could be observed that there were multiple entries for the same book because it had been misspelled.

Regardless of how difficult the data set size may make it to interpret, taking a second to realize the scope at which this data was collected is amazing and super cool. Lastly, in terms of ethical questions in regard to this data, I see no problem. The nature of this data being absent of any human actions removes any ethical concerns.

### First Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, warning = FALSE, code = readLines("chart1_comparingcheckouts.R")}
```

### Second Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("chart2_top10subjects.R")}
```

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("chart3_top10creators.R")}
```


